---
title: "DT Scrum cycle Exercise"
author: "Phillip Llewellyn"
date: "7/19/2021"
output: pdf_document
---

# Decision Tree Scrum Cycle 7

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#### Decision Tree
library(rpart)
library(rpart.plot)
library(pROC)
library(caTools)

#install.packages("RWeka")
#require(RWeka)
require(party)
require(partykit)
library(ggplot2)
```


## Step 1
- Load data and get summaries 

```{r}
data <-read.csv("BankLoan Dataset 2021 - Clean .csv")  #HR.csv
#str(data)
data$lead <- as.factor(data$lead)
data$won <- as.factor(data$won)
summary(data)
#str(data)
```


## Step 2
- Split data into training and testing data

```{r}
# separate the data for an equal split
data_lead_1 = data[data$lead == 1,]
data_lead_0 = data[data$lead == 0,]

#randomize the sampling
set.seed(1)
newDataset_lead_1 <-sample.split(Y=data_lead_1$lead, SplitRatio = 0.28)
newDataset_lead_0 <-sample.split(Y=data_lead_0$lead, SplitRatio = 0.6)

trainData <-rbind(data_lead_1[newDataset_lead_1,], data_lead_0[newDataset_lead_0,])
testData <- rbind(data_lead_1[!newDataset_lead_1,][1:150,], data_lead_0[!newDataset_lead_0,])

summary(trainData$lead)
summary(testData$lead)
```


## Step 3
- Fit a Decision Tree using training data

```{r}
# The . specifies all other columns ( Class ~ . )
DTmodel <- rpart(lead ~ education + age + job + marital + deposit + balance + loan + housing, method="class", data=trainData, parms = list (split ="information gain"), control = rpart.control(minsplit = 10, maxdepth = 5))  
```


- Target Variable = Class, 
- Input Vaiables = All, 
- split =gini or information gain 
- control = rpart.control for prepruning DT minsplit- min records at node for split to occur, maxdepth - depth of the DT

-  Fitting the model

```{r}
rpart.plot(DTmodel, type=3, extra = 2, fallen.leaves = F, cex = 0.8)
#try extra with 2,8,4, 101
```


- Print out the information

```{r}
#(DTmodel) # detailed summary of splits
DTmodel #prints the rules
```


- Run the second model

```{r}
#DTmodel2 <- J48(as.factor(Class) ~., trainData, control = Weka_control(R = TRUE, M = round(NROW(trainData)*0.05)))

#DTmodel2 <- J48(as.factor(left) ~., trainData, control = Weka_control(R = TRUE, M = 50))

#IGDT5model <- J48(as.factor(eReader_Adoption)~., trainData ,control = Weka_control(R = TRUE, M = round(NROW(trainData)*0.05))) 

#IGDT10model <- J48(as.factor(eReader_Adoption)~., trainData ,control = Weka_control(R = TRUE, M = round(NROW(trainData)*0.10)))
```


- Plot the model

```{r}
#plot(DTmodel)
```


## Step 4
- Use the fitted model to do predictions for the test data

```{r}
predTest <- predict(DTmodel, testData, type="class")
probTest <- predict(DTmodel, testData, type="prob")

actualTest <- testData$won
```



## Step 5
- Create Confusion Matrix and compute the misclassification error
 
```{r}
t1 <- table(predictions=predTest, actual = actualTest)
t1 # Confusion matrix
accuracy1 <- sum(diag(t1))/sum(t1)
accuracy1 
```


- Visualization of probabilities

```{r}
hist(probTest[,2], breaks = 100)
```


- ROC and Area Under the Curve

```{r}
ROC <- roc(actualTest, probTest[,2])
plot(ROC, col="blue")
AUC <- auc(ROC)
AUC
```


- A new dataframe with Predicted Prob, Actual Value and Predicted Value

```{r}
predicted_data <- data.frame(Probs = probTest, Actual_Value= actualTest ,Predicted_Value = predTest )  #Create data frame with prob and predictions
  #predicted_data$Probs.0 <- Class 0 Probability
  #predicted_data$Probs.1 <-  Class 1 Probability
predicted_data <- predicted_data[order(predicted_data$Probs.1, decreasing=TRUE),] # Sort on Probabilities
predicted_data$Rank <- 1:nrow(predicted_data) # Add a new variable rank
```


- plot the graph

```{r}
ggplot(data=predicted_data, aes(x=Rank, y=Probs.1)) + 
  geom_point(aes(color = Actual_Value)) + xlab("Index") + ylab("Predicted Probability of getting Diabetes")
```


## Step 6
- Use model to make predictions on newdata. Note we can specify the newData as data.frame with one or many records

```{r}
#newData <- data.frame(Nbr_Preg = 4 , Glucose_test = 100, Triceps_SF=40,BP =95, S_insulin = 150, BMI= 30,Diabetes_fn = 0.54,Age = 50 )

#predProbability <-predict(DTmodel, newData, type='prob')
#predProbability

## Performnce measures - 
#setseed(1), gini
# Simplicity = 15 leaves
# Accuracy = 0.734
# AUC = 0.7627

#setseed(1), information
# Simplicity = 10 leaves
# Accuracy = 0.71
# AUC = 0.7834

```

## Step 7
- EXAMINING STABILITY - Creating Decile Plots for Class 1 or 0 Sort

```{r}
#-----Create empty df-------
#decileDF<- data.frame(matrix(ncol=3,nrow = 0))
#colnames(decileDF)<- c("Decile","per_correct_preds","No_correct_Preds","cum_preds")
#-----Initialize varables
#num_of_deciles=10
#Obs_per_decile<-nrow(predicted_data)/num_of_deciles
#decile_count=1
#start=1
#stop=(start-1) + Obs_per_decile
#prev_cum_pred<-0
#x=0
#-----Loop through DF and create deciles
#while (x < nrow(predicted_data)) {
#  subset<-predicted_data[c(start:stop),]
#  correct_count<- ifelse(subset$Actual_Value==subset$Predicted_Value,1,0)
#  no_correct_Preds<-sum(correct_count,na.rm = TRUE)
#  per_correct_Preds<-(no_correct_Preds/Obs_per_decile)*100
#  cum_preds<-no_correct_Preds+prev_cum_pred
#  addRow<-data.frame("Decile"=decile_count,"per_correct_preds"=per_correct_Preds,"No_correct_Preds"=no_correct_Preds,"cum_preds"=cum_preds)
#  decileDF<-rbind(decileDF,addRow)
#  prev_cum_pred<-prev_cum_pred+no_correct_Preds
#  start<-stop+1
#  stop=(start-1) + Obs_per_decile
#  x<-x+Obs_per_decile
#  decile_count<-decile_count+1
#}
#------Stability plot (correct preds per decile)
#plot(decileDF$Decile,decileDF$per_correct_preds,type = "l",xlab = "Decile",ylab = "Percentage of correct predictions",main="Stability Plot for Class 1")
```















